---
title: "Untitled"
author: "Zsiros, Gabriella"
date: "2022-11-26"
output: pdf_document
---

### Prep

```{r}
library(tidyverse)
library(modelsummary)
library(stargazer)
library(fastDummies)
df <- read.csv('/users/Gabi/Downloads/morg-2014-emp.csv')
```

#### About the data

Dataset is available at <https://osf.io/g8p9j/> . The purpose of this report is to analyse earnings of men and women in a certain occupational sector.

I calculated the hourly earnings as well as its logarithmic values to help with further analysis.

```{r}
df <- df %>% 
  mutate(w = earnwke / uhours) %>%
  mutate(lnw = log(w)) 
```

#### Which occupation to choose? 
I considered that I should have approximately same amount of male data as female, and should have originally more than 500 observations per sex. Based on a short check I have selected the category of *Marketing and sales managers*.

```{r echo= FALSE}
dummies <- df %>% select(occ2012,sex) %>% dummy_cols(select_columns = 'sex')
dummies <-  dummies %>% group_by(occ2012) %>%  
  summarise('Sex1'= sum(sex_1) , 'Sex2' = sum(sex_2), 'ratio' = sum(sex_1)/sum(sex_2))

dummies %>% filter(ratio < 1.1 & ratio > 0.9 & Sex1 > 500)
```

Removing extreme values 

```{r include=FALSE}
sort(df$earnwke) %>%  head()
df <- df[order(df$earnwke),] 
df <- df[-c(1),]
df <- df[order(df$X),] 
```


#### How many hours?
A quick check of the distribution of hours has led me ot narrow it down between 20 and 60 hours per week. 

```{r include = FALSE}
hist(df$uhours)
summary(df$uhours)
```



```{r}

df <-  df %>% filter(uhours > 20 & uhours < 60) %>% filter(occ2012 == 50)
df <- df %>%  mutate(df,
                     sex_factor = factor(df$sex,labels = c('male','female')),
                     .after = sex)
df <- df %>% filter(grade92 >38)
```


### Hourly earning of men and women

#### Statistical summary

```{r, echo= F}
Range <- function(x){max(x, na.rm = T) - min(x, na.rm = T)}
P5 <- function(x){quantile(x,probs = 0.05)}
P95 <- function(x){quantile(x,probs = 0.95)}
```

```{r}
datasummary (earnwke * sex_factor ~
               Mean + Median + Min + Max + P5 + P95 + Range,
             data = df)
datasummary (w * sex_factor ~
               Mean + Median + Min + Max + P5 + P95 + Range,
             data = df)
```

It is visible in he summary that both the mean and median show difference between the two sexes. 


#### Visualizing the wage gap

```{r}
ggplot(data = df, aes(x = sex_factor, y= w, color = sex_factor))+
  geom_boxplot() + 
  scale_x_discrete(limits=rev)+
  labs(x = '', y = '$', title = "Hourly wage", ) +
  theme_classic() +
  theme(legend.position="none")
```

#### T-test

```{r}
df50f <- df %>% filter(sex == 2)
df50m <- df %>% filter(sex == 1)
t.test(df50m$w,df50f$w, mu = 0)
```

T test with value 8.5176 shows with a p -value of 2.2e-16 (very close to zero) that there is a significant difference in the average earning between men and women. Men earn 6.25-10.46 \$ more on a weekly
basis with 95% CI.

#### Linear regression

```{r}
reg1 <- lm( w ~ sex, df) 
reg2 <- lm(lnw ~ sex, df) 
huxreg('wage' = reg1,'ln wage' = reg2)

```

Applying simple regression analysis shows that women earn $8.5, i.e. 29% less on average on a weekly basis

### Introducing grade variable

#### Scatter plot with regression

```{r message = FALSE}
ggplot(data = df, aes(x = grade92, y=w, color = sex_factor))+
  geom_point()+ 
  geom_smooth(method = 'lm')+
  labs(x = '', y = '$', title = "Hourly wage", ) +
  facet_wrap(~sex_factor)+
  theme_classic() +
  theme(legend.position="none")
```

#### Summary plot wuth regression

```{r, warning= FALSE, message = FALSE}
ggplot(data = df, aes(x = grade92, y=w, color = sex_factor))+
  stat_summary()+ 
  scale_x_continuous(breaks = c(39:46))+
  geom_smooth(method = 'lm') +
  labs(x = '', y = '$', title = "Hourly wage", ) +
  xlim(39,46)+
  theme_classic()
```

#### Loess 

```{r warning= FALSE, message = FALSE}
ggplot(data = df, aes(x = grade92, y=w, color = sex_factor))+
  stat_summary()+ 
  geom_smooth(method = 'loess') +
  scale_x_continuous(breaks = c(39:46))+
  labs(x = '', y = '$', title = "Hourly wage + loess", ) +
  theme_classic()
```

Loess method in this case does not seem to be sensible, as the grade variable is a factor, rather than a numerical value.


#### Multivariate regression

```{r}
reg4 <- lm( w ~ sex + grade92, df)
reg5 <- lm( lnw ~ sex + grade92, df)
reg6 <- lm_robust(lnw ~ sex + grade92, data = df, se_type = "HC1")
huxreg('wage'=reg4,'ln wage'= reg5,'ln wage robust' = reg6)
```


Log-level transformation seems to be a more acrruate model, with lower SE-s, and higher R2. 
In this case robust SE does not show great decrease of SE, so the seond model (`reg5`) will be used to final summary.  



We can see a greater statistical significance in Bachelor's and Master's degree
### Summary
```{r}
reg7 <- lm( grade92 ~ sex, df)
huxreg('ln wage' = reg2, 'ln wage' = reg5, 'grade' = reg7,  statistics = c(N = "nobs", R2 = "r.squared"),error_pos = 'same')
```
Comparing men and women in *Marketing and Sales manager* occupational sector, analysis shows an approximate 30% difference in average salaries, considering a 20-60 work week. The second model introduces the education level, where comparing men and women in the same education level, we get a 26.8 log point difference, which here I will interpret as 27%. 
Relation between grade and sex is not to be interpreted in this case, since the education level is a factor. 