---
title: "Term Project"
author: "Zsiros, Gabriella"
date: "2022-12-20"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
```

```{r}
library(tidyverse)
library(modelsummary)
library(data.table)
library(estimatr)
library(lspline)
library(kableExtra)
```

## Introduction

As the purpose of my Analysis, I wanted to look at migration patterns of
various economies over the world. The main source I used is part of the
Worldbank Databank, where I selected some of the indicators I was
interested in. Source:
<https://databank.worldbank.org/indicator/SP.DYN.LE00.IN/1ff4a498/Popular-Indicators>

Each observation represents a country, and I had several year's data at
my disposal. As the main indicator is migration data, I wanted to find
out if there is a relationship between migration numbers and "how well"
county's economy is doing. Fo this purpose i chose one of the basic
indicator of an economy: GDP per capita. In the next paragraphs I will
introduce the dataset I specifically selected form Worldbank and will
try to determine a relationship between the variables by running a
linear regression.

## About the data:

The dependent variable (y) in question is migration, which is called
'Net migration' as indicator. This "the number of immigrants minus the
number of emigrants" that is observed during a five year period. This
means that data of 2017 actually covers the period of 2012-2017; the
data series before that, from 2007-2012.. etc. Considering the net
migration data, it is an important conclusion that observations
countries) with negative net migration value experience more residents
moving from the country than moving into. Positive numbers mean the
migration pattern just the other was around, with more people
immigrating to a country than emigration from it.

The independent variable in my dataset is GDP per Capita, in US dollar
currency.

Since the migration numbers are not really conclusive in themselves, it
its important to consider the Population of the country as well.

Aside from these variables above, and the country names, I have included
Life expectancy at birth and Human Capital index in the dataset. Life
expectancy in this context means *"Life expectancy at birth indicates
the number of years a newborn infant would live if prevailing patterns
of mortality at the time of its birth were to stay the same throughout
its life."* while Human Capital index (HCI) is a calculated value
between 0 and 1 which indicates the contributions of health and
education to worker productivity.

To make the indicators easier to interpret, I have introduced the
following naming conventions:

-   GDP per Capita is denoted as `GDPPC`

-   Life expectancy at birth denoted as `LEaB`

-   Human Capital Index as `HCI`

-   Migration with regards to the population is added as a new
    calculated variable and denoted as `Mig_rate`

For the purpose of the analysis, I selected only **one period of time:**
`2017`.

```{r}
#import dataset from csv
df <- read.csv('https://raw.githubusercontent.com/gabizsiros/DA2/main/worldbank_pop.csv')
#selecting analyzed time section
df17 <- df %>% filter(Time == 2017)
#df12 <- df %>% filter(Time == 2012)

```

```{r}
#rename important columns for readability 
df17 <- df17 %>%  rename("GDPPC" = GDP.per.capita..constant.2015.US....NY.GDP.PCAP.KD.,
                     "Migration" = Net.migration..SM.POP.NETM.,
                     "LEaB" = Life.expectancy.at.birth..total..years...SP.DYN.LE00.IN.,,
                     "Population" = Population..total..SP.POP.TOTL.,
                     "HCI" = Human.capital.index..HCI...scale.0.1...HD.HCI.OVRL.)



```

## Models

Concentrating on the renamed and relevant columns, I have decided to
remove the missing variables and keep extreme values. I have also added
a calculated column of Migration numbers divided by population, thus
creation a net migration ratio, which fares better with the intended
independent variable, `GDPPC`.

```{r}
#converting to number
df17$Migration <- as.numeric(df17$Migration)
df17$Population <- as.numeric(df17$Population)
df17$GDPPC <- as.numeric(df17$GDPPC)
df17$LEaB <- as.numeric(df17$LEaB)
df17$HCI <- as.numeric(df17$HCI)

#dropping unnecessary columns
df17 <- select(df17, -Time.Code) 

#removing missing values
df17 <- na.omit(df17, Migration,GDPPC,GDPPPP,LEaB)

#introducing migration rate
df17$Mig_rate <- df17$Migration / df17$Population


```

An overview of the data shows the `GDPPC` having a log-normal or
possibly pareto distribution, with a long right tail.

```{r}
#overview of data
datasummary_skim(df17, output = 'kableExtra', title = "Statistucal features of the dataset") %>% kable_styling(latex_options = "scale_down") %>% kable_styling(latex_options = "HOLD_position")

```

### Lowess

Comparing several models and plotting the dependent and independent
variables, a sensible choice was to take the logarithmic value of the
`GDPPC` (x) as the change/increase is better represented by approximate
percentages, rather than actual dollar values. With level x, the
observations are clustered around relatively small numbers. Following
the curve of the lowess model with level variables, the slope of the
curve intuitively changes at specific values the x is taking, namely
5000 and 25000 \$ GDP per capita which will be used late on as nodes in
a spline transformation in the analysis. *(see table in appendix)*


Taking a logarithmic value of the dependent variable (y) is not
preferred here, as the Net Migration rate takes on values that are not
positive as well.

After a log transformation of the x, Locally Weighted Scatterplot
smoothing (LOWESS) seems to capture a good curved fit, which is depicted
by default with a 95% percent confidence interval. Standard error is
visibly better where we have more data, where x takes on small values
and increases as the datapoint become more scarce.

### Linear models

After comparing more linear models the best fit probed to be the Spline
model with nodes at 5000,25000 and 45000 respectively.

```{r}
model1 <-  lm(Mig_rate ~ lspline(GDPPC,c(5000,25000,45000)), df17)

modelsummary(model1, shape = model + statistic ~ term, coef_rename = 
               c('lspline(GDPPC, c(5000, 25000, 45000))1' = '<5k',
             'lspline(GDPPC, c(5000, 25000, 45000))2' = '>5k,<25k', 
             'lspline(GDPPC, c(5000, 25000, 45000))3' = '>25k,<45k',
             'lspline(GDPPC, c(5000, 25000, 45000))4' = '>45k'),
             gof_omit = "IC|Log|F|RMSE",vcov = 'HC3', 
             output = 'kableExtra', title = 'Spline model') %>% 
  kable_styling(latex_options = "scale_down") %>% 
  kable_styling(latex_options = "HOLD_position")

```

The intercept shows that the Net migration rate where the x takes on the
value of 0 is -0.005 (meaning 0.5% of the population emigrates from the
origin country). However, since the GDP per capita's minimum value is
286.4, this information has little relevance. Under 5000, the slope
coefficient has a value of - 0.000001, as the GDP per capita increases
by 1 unit (dollar), the net migration rate decreases by 0.000001% (as
the net migration is already denoted in percentage). This can also be
expressed by 0.01% change for every 10 thousand dollar increase in GDP.
After 5000, the slope changes however, a 0.000002% increase in Net
migration rate until 25000 then 0.0000005 decrease (almost a flat line,
this is one decimal lower than the other coefficients), and finally a
0.0000008 increase. Also can be interpreted as 0.01% decrease, 0.02%
increase, 0.005% decrease, then 0.08 increase by every 10 thousand
dollars increase in the countries GDP per Capita


```{r,fig.height = 5}
#spline plot
ggplot (df17, aes(GDPPC,Mig_rate)) +
  geom_point() +
  geom_smooth(method='lm',formula =  y ~ lspline(x,c(0,5000,25000,45000,max(df17$GDPPC)))) + 
  labs (x = 'GPD per Capita', y = 'Net migration rate', title = 'Spline') +
  scale_x_continuous(breaks = c(0,5000,25000,45000,max(df17$GDPPC))) +
  xlim(0,45000) +
  theme_bw()

```

## Alternatives

As alternative models, I have analyzed two alternative models: one not
transformation the variables, only considering the Migration as ratio,
while the other took the logarithmic value of GDP per capita. Using
robust standard errors, both the R^2^ and adjusted R^2^ supports the
original Spline.

The second model (level-level) shows a 0.0000006 increase in net
migration rate by increasing the GDP Per capita on average, while the
third model (level-log) shows a 0.00007increase in the migration rate
when the GDP per capita is increased by 1%.

```{r}
model1 <-  lm(Mig_rate ~ lspline(GDPPC,c(5000,25000,45000)), df17)
model2 <- lm(Mig_rate ~ GDPPC,df17)
model3 <- lm(Mig_rate ~log(GDPPC),df17)


modelsummary(list(model1,model2,model3), 
             coef_rename = 
               c('lspline(GDPPC, c(5000, 25000, 45000))1' = '<5k',
             'lspline(GDPPC, c(5000, 25000, 45000))2' = '>5k,<25k', 
             'lspline(GDPPC, c(5000, 25000, 45000))3' = '>25k,<45k',
             'lspline(GDPPC, c(5000, 25000, 45000))4' = '>45k'),
             gof_omit = "IC|Log|F|RMSE",vcov = 'HC3', output = 'kableExtra', title = "Model comparison")  %>% kable_styling(latex_options = "HOLD_position")
#  kable_styling(latex_options = "scale_down")

```

### Multivariate regression

Using additional variable such as life Expectancy at Birth and Human Capital Index introduces more granularity to the data. Taking the nominal value of GDP per capita, the model shows that in countires with the same GDP per capita level, the migration increases be 0.0005% with a higher life expectancy, while the rate decreases by 0.042% the the HCI is increased and all aohter variables remain the same.

```{r}

model4 <- lm(Mig_rate ~ GDPPC + LEaB + HCI,df17)


modelsummary(model4, shape = model + statistic ~ term, gof_omit = "IC|Log|F|RMSE",vcov = 'HC3', output = 'kableExtra', title = "Multivariate model")  %>%
 kable_styling(latex_options = "HOLD_position")

```

## Summary and conclusion

The regression of the dataset could be captured with various methods, of which the spline model proved to be the most accurate. Depending on the the GDP per capita range, the change in net migration rate varies, at some point more significant than others, hence the fit of the spline model. 
Introducing more variable from the several indicator available at WorldBank, we can introduce some granularity which opens to topic for further discussions on what makes a country a migration target, on top the the GDP economic indicator. 

## Appendix

### Level-level plot with lowess
```{r, fig.height = 4}
ggplot (df17, aes(GDPPC,Migration)) +
  geom_point() +
  geom_smooth(method = loess) +
  labs (x = 'GPD per Capita', y = 'Net migration rate', title = 'Level-Level LOWESS') +
  theme_bw() 
```

### Level-Log plot
```{r, fig.height = 3}
plot1 <- ggplot (df17, aes(log(GDPPC),Mig_rate), color = method) +
  geom_point() +
  geom_smooth(method = lm, color = 'blue') + 
  labs (x = 'ln(GPD per Capita)', y = 'Net migration rate', title = 'Level-Log') +
  theme_bw() 

plot1 + theme(plot.title = element_text(hjust = 0.5))
```

### Level-level plot
```{r, fig.height = 4}
ggplot (df17, aes(GDPPC,Migration)) +
  geom_point() +
  geom_smooth(method = lm) +
  labs (x = 'GPD per Capita', y = 'Net migration rate', title = 'Level-Level') +
  theme_bw() 
```



### Datatable
```{r}
head(df17 %>% select(Country.Name,Migration,Population,GDPPC, LEaB,HCI),20)
```
